The Sub-Millimetre User Reduction Facility: an iterative map-maker for \scuba

### 2012-08-22 smurf_paper.tex
updates to most of the paper from Douglas Scott

\begin{abstract}
  We describe data properties of the Submillimetre Common User
  Bolometer Array 2 (\scuba), a camera for the 15-m James Clerk
  Maxwell Telescope that provides simultaneous imaging capabilities at
  450 and 850\,\micron, and the production of maps using the
  Sub-Millimetre User Reduction Facility (SMURF). With nearly 5000
  bolometers in each imaging band, and an approximately 5\,arcmin
  field-of-view, \scuba\ is presently the largest submillimetre camera
  in the world, and produces data at rates an order-of-magnitude
  faster than previous instruments. We have adopted a fast, iterative
  approach to map-making that enables data reduction on single,
  modern, high-end desktop computers with execution times that are
  shorter than the observing times.  SMURF is used in an automated
  setting both at the telescope for real-time feedback to observers,
  as well as for the production of generic science products at the
  Canadian Astronomical Data Centre. A suite of tools are also
  provided for interactive data analysis.  Three detailed case studies
  are used to: (i) explore convergence properties of the map-maker
  using simple prior constraints (Uranus -- a point source); (ii)
  demonstrate that our strategy is capable of recovering angular
  scales up to the size of the array footprint (approximately
  5\,arcmin) for bright extended sources (star-forming region M17);
  and (iii) achieve the white-noise limit of the instrument for faint
  point-source studies (blank-field survey of the Lockman Hole).
\end{abstract}

@@

The Submillimetre Common User Bolometer Array 2
\citep[\scuba,][]{2006SPIE.6275E..45H,holland2012} is a new instrument for the 15-m James
Clerk Maxwell Telescope (JCMT) on Mauna Kea, Hawai'i. The camera
simultaneously images the sky in two broad bands centered over 450 and
850\,\micron, with approximately 7.5 and 14.5\,arcsec full-width at
half-maximum (FWHM) point spread functions (PSFs). The focal planes at
each wavelength are populated with 4 rectangular subarrays, consisting
of $40 \times 32$ bolometers each, and subtend a nearly contiguous $7'
\times 7'$ footprint on the sky.  Ultimately, the instrument will also
be equipped with a Fourier transform spectrometer
\citep[FTS-2,][]{gom2010}, and a polarimeter
\cite[POL-2,][]{2011ASPC..449...68B}. This paper describes the properties of
\scuba\ data that are relevant to producing maps of the imaging data,
and the Submillimetre User Reduction Facility, SMURF, a software
package for perfoming the reduction written using the Starlink
Software Environment \citep{2009ASPC..411..418J}. The analysis
of FTS-2 and POL-2 data will be described at a later date once these additional
instruments have been commissioned. The details of the instrument
design, performance, and its calibration are given in two companion
papers: \citet{holland2012} and \citet{dempsey2012}.

Over the last twenty years, observations in the submillimetre band
(defined here to be 200--1200\,\micron) have helped revolutionize
several important areas of astrophysics, including: the
characterization of early stages of star-formation by identifying the
dense, cold regions in molecular clouds where stars may eventually
form; discovering through blind surveys a class of massive dusty
star-forming galaxies in the early ($z>2$) Universe, now referred to
as submillimetre galaxies, or SMGs; and identifying debris disks
around nearby stars, the early stages of planet formation.  With
$\sim10000$ total detectors, \scuba\ is presently the largest
submillimetre camera in the world, and the fastest wide-area
ground-based submillimetre imager.

Submillimetre imaging cameras generally use bolometers to maximize
sensitivity, in which an absorber collects photons that pass through
broad-band filters, and the signal is read out-out using a thermometer
\citep[e.g.][]{griffin2002}. The sensitivity of bolometers in a camera
like \scuba\ are limited by white photon and phonon noise from the
instrument and ambient backgrounds. The low-frequency noise, however,
is dominated by sources such as slow variations in mean backgrounds
(e.g., thermal variations within the cryostat, and the atmosphere for
ground-based cameras), and drift in the readout electronics. Such
noise has a power spectrum $\propto 1/f^\alpha$ ($\alpha>0$), and the
frequency at which it is comparable to the white noise level is called
the ``$1/f$ knee''. Since the low-frequency noise is largely
correlated between all, or subsets, of the bolometers in time, it can
be suppressed during map-making, since astronomical signals have the
distinct property that they are fixed in a sky reference frame
(assuming they are not time-varying). If successful, the noise in the
resulting map is said to be ``white noise limited'', meaning that the
map noise is uncorrelated spatially, and has an amplitude that scales
as the $\mathrm{NEFD}/\sqrt{t}$, where the NEFD is the
noise-equivalent flux-density (the white noise level of a bolometer in
1\,s of integration), and $t$ is the amount of integration time in a
map pixel.

There are numerous ways to attack the problem of map production, both
in terms of the data-collection method, and processing. The two most
important principles to follow in terms of scan strategy are: (i) to
modulate the astronomical signals of interest in such a way that they
appear in the lowest-noise regions of the bolometer noise power
spectrum, i.e., above the $1/f$ knee; and (ii) to provide good
``cross-linking'', in which each portion of the map is sampled on a
range of temporal scales, again, to help distinguish time-varying
noise features from fixed astronomical sources. In the case of \scuba,
(i) is achieved through fast-scanning of the entire telescope (up to
600\,arcsec\,sec$^{-1}$ slew speeds), such that significant drift in
the bolometers due to low-frequency noise occurs more slowly than the
crossing times for the astronomical scales of interest; and (ii) by
offering scan patterns that cross the sky at a wide range of position
angles, ultimately observing each portion of the map at a number of
different times. Such methods are now used by virtually all existing
ground-based submillimetre cameras, in favour of ``chopping'' methods
(where the secondary is quickly moved to modulate the signal) that
were more appropriate for older instruments that had poorer
low-frequency noise performance, and were only sensivitve to modest
angular scales
\citep[e.g.][]{glenn1998,weferling2002,wilson2008,kovacs2008b}.

There are three general styles of map-making that are relevant to
reducing bolomter data in the literature. The simplest ``direct
methods'' involve some basic processing of the data to remove as much
noise contamination as possible, e.g., using baseline removal,
filters, and then simply re-gridding the data into a map. Such was the
basic recipe for the baseline reduction of chopped data from SCUBA-2's
predecessor SCUBA \citep{1998ASPCd..145..216J,2000ASPC..216..559J},
and MAMBO, another camera from the same generation. Generally
speaking, such methods are fast, although dependending on the science
goals and noise properties of the data,
they may not achieve the best theoretical noise performance on the
angular scales of interest. A particular case of this method was
successfully used for the reduction of data from BOLOCAM
\citep[e.g.][]{laurent2005}, and its younger sibling AzTEC
\citep[e.g.][]{scott2008}, using ``principal component analysis''
(PCA) cleaning. A statistical black-box is used to remove the most
correlated components of the bolometer signals, enabling the detection
of point-sources very close to the theoretical white-noise limits of
the detectors with reasonable computation times. However, PCA cleaning
is not a good generic solution for producing maps of extended
structures, since large astronomical sources produce correlated
signals amongst many detectors, and are removed by this procedure.

The best map-making strategies for recovering information on all
angular scales are \emph{maximum likelihood} techniques, in which the
time-series data are expressed as a sampling of the ``true'' map of
the sky plus noise, and then the equation is inverted to estimate the
map as some weighted linear combination of the data that minimizes the
variance.  The first good description of this method appears in
\citet{janssen1992} for the production of maps from the COsmic
Background Explorer (COBE -- the description is relevant despite the
fact that it used a differential radiometer instead of
bolometers). Other descriptions in the CMB literature include
\citet{tegmark1997} and \citet{stompor2002}.  The downside to such
methods is that they can be both computationally expensive and have
large memory requirements. While for some experimental designs fast
iterative methods for the inversion do exist with minimal memory usage
\citep[e.g.][]{wright1996}, for the more general map-making problem
involving many detectors, the problem is significantly complicated
both by the need to measure the cross power-spectra for all unique
pairs of bolometers, and then to perform the inversion itself. The
most promising maximum-likelihood method that may one day be applied
to SCUBA-2 data is ``SANEPIC'', which successfully reduced maps from
the Balloon-borne Large-Aperture Submillimeter Telescope, involving
hundreds of bolometers, and correctly incorporating inter-bolometer
noise correlations \citep[][]{patanchon2008}.

@@

In this section we summarize how \scuba\ bolometers detect temperature
variations (Section~\ref{sec:bolos}), typical \scuba\ scan strategies
(Section~\ref{sec:scan}).

@@

While the details of how \scuba\ works is described in
\citet{holland2012}, and its calibration in \citet{dempsey2012}, we
summarize the basic concepts relative to map-making here.

The \scuba\ bolometer temperatures are read-out using superconducting
transition-edge sensors (TESs), with critical temperatures
$T_\mathrm{C}$ (...typical value...). The range about $T_\mathrm{C}$
in which the transition from a superconductor to a normal material
occurs is extremely small (...typical value...?). If the devices are
biased to a temperature within this transition on large time scales,
then tiny temperature fluctuations on short time-scales will produce
significant changes in the impedence. The entire \scuba\ focal plane
is therefore held at a \emph{base temperature} somewhat below
$T_\mathrm{C}$. The TESs then have a small electrical voltage biases,
and individual pixel heaters (using a current bias to apply the
desired Joule heating) are used to place the detectors in the
transition. Only a small detector bias voltage is required since the
heaters compensate for most of the variations in the base temperature,
and other ambient loads (like the atmosphere).  Small temperature (and
therefore impedence) fluctuations produce a changing current through
the TESs. The resulting changing magnetic fields are then amplified
using chains of superconducting quantum interference devices (SQUIDs),
and the larger output current is finally digitized. Each detector has
its own SQUID for the first-stage of the amplification, but the
remaining stages occur within a common chain of SQUIDs for each column
of 40 detectors. Each of the 32 rows is then read-out in sequence at a
row-visit frequency of ...kHz?... This multiplexing scheme greatly
reduces the number of readout and bias lines to each detector, making
such large-format bolometer array designs feasible. It should also be
noted that there is an additional 33rd row of SQUID reasouds that are
not connected to TES. These ``dark SQUIDs'' are used to track
non-thermal noise sources that are common to each column's amplifier
chain.

The response of the output digitized current, $I$, with respect to a
change in the power $P$ incident on the bolometer, $dI/dP$ depends on
the efficiency of the thermal absorber, how well the TES is biased
into the transition, and the gain of the SQUID amplifiers. Since there
are variations in $T_\mathrm{C}$ across each subarray due to
manufacturing difficulties, considerable effort has gone into
dynamically applying position-dependent biases as well as an over-all
focal plane bias to make as many detectors useable as possible
\citep{holland2012}. In practice, the response of a given detector can
therefore vary somewhat from observation to observation depending on
how the bias and amplification parameters were set up. To track these
responsivity changes, at the start of each observation, the current is
ramped through the bolometer heaters such that $dI/dP$ can be
measured. Note that the bolometers may have a slightly different
response to heater power due to conductance as compared to the
response to optical power, so additioanl efficiency corrections must
also be obtained. While not discussed further in this paper, details
may be found in \citet{dempsey2012}. We will refer to measurements of
$dI/dP$ across the focal plane as the
\emph{flatfield}. Fig.~\ref{fig:sensitivities} shows the flatfield for
each focal plane.

@@

In general, bolometers have two predominant noise components: (i) the
effectively white thermal noise produced by the phonon and optical
loading on the absorbers; and (ii) and longer-timescale drift
resulting from slow variations in the mean load (e.g. fridge,
telescope, or atmospheric changes), and also the readout
electronics. As such, the spectrum corresponding to a bolometer
time-series exhibits a ``$1/f$'' shape below some knee frequency,
$f_\mathrm{knee}$, and is relatively flat at higher frequencies. Other
smaller effects may include line features in the spectrum (a common
culprit is pickup from AC mains, or microphonic signals induced by
mechanical vibrations), and finally a roll-off towards the Nyquist
frequency caused by the anti-aliasing filter that precedes
digitization.

The final conversion of power received by the detector into
astronomical flux units requires two more fundamental corrections: (i)
atmospheric extinction (due to atmospheric water vapour) must be
compensated for; and (ii) the corrected power is related to an
astronomical flux density through the effective area of the detector,
the spectral response of the bolometers (bandpass), the spectrum of
the source, and additional optical efficiency factors -- a \emph{flux
  conversion factor} (FCF). Extinction is measured, in most cases,
using the JCMT line-of-sight water vapour radiometer (WVM) which is
calibrated using skydips and the CSO 225\,GHz tipper. The FCF is
determined empirically using measurements of standard sources. For
further details on these calibration steps, see \citet{dempsey2012}.

@@

In order to observe astronomical sources, it is clearly advantageous
that the signals appear in the region of the spectrum that offers the
highest \snr, namely frequencies above $f_\mathrm{knee}$. This effect
is achieved by modulating the signal; the original SCUBA primarily
used the JCMT chopping secondary to place sources into the spectrum at
about $\sim$10\,Hz. The problem with this strategy was that SCUBA was
ultimately limited to recovering information on angular scales smaller
than the chop throw (a maximum of a couple of arcmin)... cite
somebody. The primary approach to map-making with \scuba\ is instead
to slew the entire telescope at such a rate that angular scales of
interest are crossed on time scales shorter than $1/f_\mathrm{knee}$,
similar to other ground-based instruments (MAMBO, BOLOCAM, SHARC-II,
AzTEC, LABOCA,...). The desire to scan rapidly has two basic
constraints: (i) the telescope cannot physically move faster that
several hundred arcsec\,sec$^{-1}$ without requiring unacceptably
large accelerations to change direction, and (ii) bolometers are
designed with a thermal time constant that balance response with
sensitivity. Taking these constraints into consideration, \scuba\ was
ultimately designed to work at a maximum scan velocity of
600\,arcsec\,sec$^{-1}$. The data acquisition system has a target
sample rate of 200\,Hz, which at that scan velocity results in a
sample every 3.0\,arcsec, or approximately one third of the
450\,\micron\ diffraction-limited full-width at half-maximum (FWHM) --
a typical rule-of-thumb for adequately sampling a Gaussian point
spread function (PSF). Similarly, the \scuba\ bolometers are designed
to have a time constant of 2 ms \textbf{check that number!} in order
to avoid distorting PSF-sized sources at this maximum scan rate.

The scan strategy is designed, primarily, to mitigate the effects of
$1/f$ drift. The simplest scan one might use is a raster pattern, or a
series of parallel passes across this field. However, such a scan
yields extremely poor performace for structures that are transverse to
the scan direction, since a long time will have ellapsed between each
parallel step, and the signal will therefore be totally dominated by
the $1/f$ noise. In addition, other noise features may appear in the
data periodically. With such a regularly spaced periodic scan
strategy, these noise features can easily conspire with the scan
pattern to produce similarly periodic noise structures in the map, or
\emph{scan-synchronous} noise. It has therefore been recognized that
scan strategies that visit each region of the map on as many different
time scales, and scan angles as possible offer the best protection
against large-scale noise (...cite some people...); such strategies
are said to offer good \emph{cross-linking}. As a tradeoff between
truly random scan patterns and feasibility of implementation taking
into consideration both mechanical and scheduling constraints, \scuba
presently has three primary scan modes: PONG (cite Borys et al.),
Lissajous, and Daisy... (more citations? Kovacs et al. has a nice
description... also CMB literature).

@@

Both bolometers share significant long-timescale structure
($\gsim10$\,s) that appears to be related to variations in the fridge
base temperature, although the similarity is clearly greater at
850\,\micron. In this particular case, the total power in the
fluctuations at 450\,\micron, from $-0.04$\,pW to $+0.08$, are larger
than the $-0.03$\,pW to $+0.05$\,pW fluctuations at 850\,\micron. Such
behaviour might be expected if there is a comparable varying thermal
load from the fridge at each wavelength, but a larger contribution of
atmospheric variations through the 450\,\micron\ bandpass filters. We
also note that there is no obvious strong correlation between the
low-frequency signal structure, at either wavelength, with the
telescope motion.

The low-frequency signal component of the bolometer signals is also
highly correlated amongst bolometers in the same subarray. We have
calculated a common-mode signal, $\mathbf{c}(t)$, as the average time
series of all the working bolometers. We then fit the amplitude of
$\mathbf{c}(t)$ at each wavelength to the signals shown in
Fig.~\ref{fig:bolos_mix} and remove it, yielding the grey residual
signals. These residuals are quite flat, although still with
noticeable variations. The white noise is also apparent, and larger at
450\,\micron\ as one would expect from the larger backgrounds compared
to 850\,\micron.

Next, we produce power spectral density (PSD) plots for several
representative bolometers in Fig.~\ref{fig:pspec}. To produce this
figure, we follow the convention that the PSD as a function of
frequency, $\mathbf{P}(f)$, is normalized such that the integral over
frequency gives the same variance as the time-series variance across
the full time series. In other words, given a bolometer signal
$\mathbf{b}(t)$,
%
\begin{equation}
\label{eq:psd}
\langle\mathbf{b}^2(t)\rangle = \int \mathbf{P}(f) df .
\end{equation}
%
However, we only show the PSD up to the Nyquist frequency, so the
conversion from variance measured in the frequency domain to the time
domain requires and additional multiplcation by 2; hence the units are
pW$^2$\,Hz$^{-1}$ rather than pW$^2$\,s.

>>>>

\begin{abstract}
  The Submillimetre Common User Bolometer Array 2 (\scuba) is an
  instrument operating on the 15-m James Clerk Maxwell Telescope,
  consisting of about 5000 bolometers in each of two simultaneous
  imaging bands centered over 450 and 850\,\micron. The camera is
  scanned across the sky and records data at a rate of 200\,Hz. This
  data volume is much greater than previously existing submillimetre
  cameras, and represents a significant data anlaysis challenge. We
  describe the production of SCUBA-2 maps using the Sub-Millimetre
  User Reduction Facility (SMURF) in which we have adopted a fast,
  iterative approach to map-making that enables data reduction on
  single, modern, high-end desktop computers with execution times that
  are shorter than the observing times.  SMURF is used in an automated
  setting both at the telescope for real-time feedback to observers,
  as well as for the production of generic science products at the
  Canadian Astronomical Data Centre. A suite of tools are also
  provided for interactive data analysis.  Three detailed case studies
  are used to: (i) explore convergence properties of the map-maker
  using simple prior constraints (Uranus -- a point source); (ii)
  achieve the white-noise limit of the instrument for faint
  point-source studies (blank-field survey of the Lockman Hole); and
  (iii) demonstrate that our strategy is capable of recovering angular
  scales up to the size of the array footprint (approximately
  5\,arcmin) for bright extended sources (star-forming region M17).
\end{abstract}

@@

The Submillimetre Common User Bolometer Array 2
\citep[\scuba,][]{2006SPIE.6275E..45H,holland2012} is a new instrument
for the 15-m James Clerk Maxwell Telescope (JCMT) on Mauna Kea,
Hawai'i. The camera simultaneously images the sky in two broad bands
centered over 450 and 850\,\micron, with approximately 7.5 and
14.5\,arcsec full-width at half-maximum (FWHM) point spread functions
(PSFs). The focal planes at each wavelength are populated with 4
rectangular subarrays, consisting of $40 \times 32$ bolometers each,
and subtend a nearly contiguous 7\,arcmin $\times$ 7\,arcmin footprint
on the sky.  The instrument is also equipped with a Fourier transform
spectrometer \citep[FTS-2,][]{gom2010}, and a polarimeter
\cite[POL-2,][]{2011ASPC..449...68B}. This paper describes the
properties of \scuba\ data that are relevant to producing maps of the
imaging data, and the Submillimetre User Reduction Facility, SMURF, a
software package for perfoming the reduction written using the
Starlink Software Environment \citep{2009ASPC..411..418J}. The
analysis of FTS-2 and POL-2 data will be described at a later date
once these additional instruments have been commissioned and
characterized. The details of the instrument design, performance, and
its calibration are given in two companion papers: \citet{holland2012}
and \citet{dempsey2012}.

Over the last twenty years, observations in the submillimetre band
(defined here to be 200--1200\,\micron) have helped revolutionise
several important areas of astrophysics, including: the
characterisation of the early stages of star-formation by identifying the
dense, cold regions in molecular clouds where stars may eventually
form; discovering through blind surveys a class of massive dusty
star-forming galaxies in the early ($z>2$) Universe, now referred to
as submillimetre galaxies, or SMGs; and identifying debris disks
around nearby stars, helping us understand the early stages of planet formation.  With
$\sim10000$ total detectors, \scuba\ is presently the largest
submillimetre camera in the world, and the fastest wide-area
ground-based submillimetre imager.

Submillimetre imaging cameras generally use bolometers to maximise
sensitivity, in which an absorber collects photons that pass through
broad-band filters, and the signal is read out using a thermometer
\citep[e.g.][]{griffin2002}. The sensitivity of bolometers in a camera
like \scuba\ are limited by white photon and phonon noise from the
instrument and ambient backgrounds. The low-frequency noise, however,
is dominated by sources which produce slow variations in mean backgrounds
(e.g., thermal variations within the cryostat, and the atmosphere for
ground-based cameras), and drifts in the readout electronics. Such
noise has a power spectrum $\propto 1/f^\alpha$ ($\alpha>0$), and the
frequency at which it is comparable to the white noise level is called
the ``$1/f$ knee''. Since the low-frequency noise is largely
correlated between all, or subsets, of the bolometers in time, it can
be suppressed during map-making, since astronomical signals have the
distinct property that they are fixed in a sky reference frame
(assuming they are not time-varying). If successful, the noise in the
resulting map is said to be ``white noise limited'', meaning that the
map noise is uncorrelated spatially, and has an amplitude that scales
as the $\mathrm{NEFD}/\sqrt{t}$, where the NEFD is the
noise-equivalent flux-density (the white noise level of a bolometer in
1\,s of integration), and $t$ is the amount of integration time in a
map pixel.

There are numerous ways to attack this map-making problem, both in
terms of the data-collection method, and processing. The two most
important principles to follow in terms of scan strategy are: (i) to
modulate the astronomical signals of interest in such a way that they
appear in the lowest-noise regions of the bolometer noise power
spectrum, i.e., above the $1/f$ knee; and (ii) to provide good
``cross-linking'', in which each portion of the map is sampled on a
range of temporal scales, again, to help distinguish time-varying
noise features from fixed astronomical sources. In the case of \scuba,
(i) is achieved through fast-scanning of the entire telescope (up to
600\,arcsec\,sec$^{-1}$ slew speeds), such that significant drift in
the bolometers due to low-frequency noise occurs more slowly than the
crossing times for the astronomical scales of interest; and (ii) by
offering scan patterns that cross the sky at a wide range of position
angles, ultimately observing each portion of the map at a number of
different times. Such methods are now used by virtually all existing
ground-based submillimetre cameras
\citep[e.g.][]{glenn1998,weferling2002,wilson2008,kovacs2008b}., in
favour of ``chopping'' methods (where the secondary is quickly moved
to modulate the signal) that were more appropriate for older
instruments that had poorer low-frequency noise performance, and were
only sensivitve to modest angular scales

There are three general styles of map-making that are relevant to
reducing bolomter data in the literature. The simplest ``direct
methods'' involve some basic processing of the data to remove as much
noise contamination as possible, (e.g., using baseline removal and
other simple filters), and then re-gridding these cleaned data into a
map. Such was the basic recipe for the baseline reduction of chopped
data from SCUBA-2's predecessor SCUBA
\citep{1998ASPC..145..216J,2000ASPC..216..559J}, and MAMBO, another
camera from the same generation. Generally speaking, such methods are
fast, although dependending on the science goals and noise properties
of the data, they may not achieve the best noise performance on the
angular scales of interest. A particular case of this method was
successfully used for the reduction of data from BOLOCAM
\citep[e.g.][]{laurent2005}, and its younger sibling AzTEC
\citep[e.g.][]{scott2008}, using ``principal component analysis''
(PCA) cleaning. A statistical black-box removes the most correlated
components of the bolometer signals, enabling the detection of
point-sources very close to the theoretical white-noise limits of the
detectors with reasonable computation times. However, PCA cleaning is
not a good generic solution for producing maps of extended structures,
since such sources produce correlated signals amongst many detectors,
and are removed by this procedure.

The best map-making strategies for recovering information on all
angular scales are \emph{maximum likelihood} techniques, in which the
time-series data are expressed as a sampling of the ``true'' map of
the sky plus noise, and then the equation is inverted to estimate the
map as some weighted linear combination of the data that minimizes the
variance.  The first good description of this method appears in
\citet{janssen1992} for the production of maps from the COsmic
Background Explorer (COBE -- the description is relevant despite the
fact that it used a differential radiometer instead of
bolometers). Other descriptions in the CMB literature include
\citet{tegmark1997} and \citet{stompor2002}, and an application to
data from SCUBA is described in \citet{borys2004}.  The downside to
such methods is that they can be both computationally expensive and
have large memory requirements. While for some experimental designs
fast iterative methods for the inversion do exist with minimal memory
usage \citep[e.g.][]{wright1996}, for the more general map-making
problem involving many detectors, the problem is significantly
complicated both by the need to measure the cross power-spectra for
all unique pairs of bolometers, and then to perform the inversion
itself. The most promising maximum-likelihood method that may one day
be applied to SCUBA-2 data is ``SANEPIC'', which successfully reduced
maps from the Balloon-borne Large-Aperture Submillimeter Telescope,
involving hundreds of bolometers, and correctly incorporating
inter-bolometer noise correlations \citep[][]{patanchon2008}.

@@

In this section we summarize how \scuba\ bolometers detect temperature
variations (Section~\ref{sec:bolos}), describe \scuba\ scan strategies
(Section~\ref{sec:scan}), typical bolometers signals
(Section~\ref{sec:bolosignal}), and the impact of magnetic field
pickup (Section~\ref{sec:magpickup}, describe the differences between
SCUBA-2 Shared Risk Observing data and those of the the fully
commissioned instrument (Section~\ref{sec:s2sro}), and finally
illustrate the use of principal component analysis
(Section~\ref{sec:pca}).

@@

While the details of how \scuba\ works are described in
\citet{holland2012}, and its calibration in \citet{dempsey2012}, we
summarize the basic concepts relevant to map-making here.

The \scuba\ bolometer temperatures are read-out using superconducting
transition-edge sensors (TESs), with critical temperatures
$T_\mathrm{C}$ (...typical value...). The range about $T_\mathrm{C}$
in which the transition from a superconductor to a normal material
occurs is extremely small (...typical value...?). If the devices are
biased to a temperature within this transition on large time scales,
then tiny temperature fluctuations on short time scales will produce
significant changes in the impedence. The entire \scuba\ focal plane
is therefore held at a \emph{base temperature} somewhat below
$T_\mathrm{C}$. The TESs then have small electrical voltage biases,
and individual pixel heaters (using a current bias to apply the
desired Joule heating) are used to place the detectors in the
transition. Only a small detector bias voltage is required since the
heaters compensate for most of the variations in the base temperature,
and other ambient loads (like the atmosphere).  Small temperature (and
therefore impedence) fluctuations produce a changing current through
the TESs. The resulting changing magnetic fields are then amplified
using chains of superconducting quantum interference devices (SQUIDs),
and the larger output current is finally digitized. Each detector has
its own SQUID for the first-stage of the amplification, but the
remaining stages occur within a common chain of SQUIDs for each column
of 40 detectors. Each of the 32 rows is then read-out in sequence at a
row-visit frequency of ...kHz?... This multiplexing scheme greatly
reduces the number of readout and bias lines to each detector, making
such large-format bolometer array designs feasible. It should also be
noted that there is an additional 33rd row of SQUID readouts that are
not connected to TES. These ``dark SQUIDs'' are used to track
non-thermal noise sources that are common to each column's amplifier
chain.

The response of the output digitized current, $I$, with respect to a
change in the power $P$ incident on the bolometer, $dI/dP$ depends on
the efficiency of the thermal absorber, how well the TES is biased
into the transition, and the gain of the SQUID amplifiers. Since there
are variations in $T_\mathrm{C}$ across each subarray due to
manufacturing differences, considerable effort has gone into
dynamically applying position-dependent biases as well as a total
focal plane bias to make as many detectors usable as possible
\citep{holland2012}. In practice, the response of a given detector can
therefore vary somewhat from observation to observation, depending on
how the bias and amplification parameters were set up. To track these
responsivity changes, at the start of each observation, the current is
ramped through the bolometer heaters such that $dI/dP$ can be
measured. Note that the bolometers may have a slightly different
response to heater power due to conductance as compared to the
response to optical power, so additioanl efficiency corrections must
also be obtained. While not discussed further in this paper, details
may be found in \citet{dempsey2012}. We will refer to measurements of
$dI/dP$ across the focal plane as the
\emph{flatfield}. Fig.~\ref{fig:sensitivities} shows the flatfield for
each focal plane.

@@

In general, bolometers have two predominant noise components: (i) the
effectively white thermal noise produced by the phonon and optical
loading on the absorbers; and (ii) and longer timescale drift
resulting from slow variations in the mean load (e.g. fridge,
telescope, or atmospheric changes), and also the readout
electronics. As such, the spectrum corresponding to a bolometer
time-series exhibits a ``$1/f$'' shape below some knee frequency,
$f_\mathrm{knee}$, and is relatively flat at higher frequencies. Other
smaller effects may include line features in the spectrum (a common
culprit is pickup from AC mains, or microphonic signals induced by
mechanical vibrations), and finally a roll-off towards the Nyquist
frequency caused by the anti-aliasing filter that precedes
digitization.

The final conversion of power received by the detector into
astronomical flux units requires two more fundamental corrections: (i)
atmospheric extinction (due to atmospheric water vapour) must be
compensated for as a function of time; and (ii) the corrected power is
related to an astronomical flux density through the effective area of
the detector, the spectral response of the bolometers (bandpass), the
spectrum of the source, and additional optical efficiency factors -- a
\emph{flux conversion factor} (FCF). Extinction is measured, in most
cases, using the JCMT line-of-sight water vapour radiometer (WVM)
which is calibrated using sky-dips and the CSO 225\,GHz tipper. The
FCF is determined empirically using measurements of standard
sources. For further details on these calibration steps, see
\citet{dempsey2012}.

@@

In order to observe astronomical sources, it is clearly advantageous
that the signals appear in the region of the spectrum that offers the
highest \snr, namely frequencies above $f_\mathrm{knee}$. This effect
is achieved by modulating the signal; the original SCUBA primarily
used the JCMT chopping secondary to place sources into the spectrum at
about $\sim$10\,Hz. The problem with this strategy was that SCUBA was
ultimately limited to recovering information on angular scales smaller
than the chop throw (a maximum of several arcmin). The primary
approach to map-making with \scuba\ is instead to slew the entire
telescope at such a rate that angular scales of interest are crossed
on time scales shorter than $1/f_\mathrm{knee}$, similar to other
ground-based instruments (MAMBO, BOLOCAM, SHARC-II, AzTEC,
LABOCA,...). The desire to scan rapidly has two basic constraints: (i)
the telescope cannot physically move faster that several hundred
arcsec\,sec$^{-1}$ without requiring unacceptably large accelerations
to change direction; and (ii) bolometers are designed with a thermal
time constant that balances response with sensitivity. Taking these
constraints into consideration, \scuba\ was ultimately designed to
work at a maximum scan speed of 600\,arcsec\,sec$^{-1}$. The data
acquisition system has a target sample rate of 200\,Hz, which at that
scan velocity results in a sample every 3.0\,arcsec, giving
approximately 2.5 samples per 450\,\micron\ diffraction-limited
full-width at half-maximum (FWHM) -- a typical rule-of-thumb for
adequately sampling a Gaussian point spread function (PSF). Similarly,
the \scuba\ bolometers are designed to have a time constant of 2 ms
\textbf{check that number!} in order to avoid distorting PSF-sized
sources at this maximum scan rate.

The scan strategy is designed, primarily, to mitigate the effects of
$1/f$ drift. The simplest scan one might use is a raster
``boustrophedon'' pattern, i.e. a series of parallel passes across the
field. However, such a scan yields extremely poor performace for
structures that are transverse to the scan direction, since a long
time will have ellapsed between each parallel step, and the signal
will therefore be totally dominated by the $1/f$ noise. In addition,
other periodic noise features may appear in the data. With such a
regularly spaced scan strategy, these noise features can easily
conspire with the scan pattern to produce similarly periodic noise
structures in the map, or \emph{scan-synchronous} noise. It has
therefore been recognized that scan strategies that visit each region
of the map on as many different time scales, and scan angles as
possible offer the best protection against large-scale noise
\citep[e.g.][]{tegmark1997,kovacs2008b}; such strategies are said to
offer good \emph{cross-linking}. As a tradeoff between truly random
scan patterns and feasibility of implementation taking into
consideration both mechanical and scheduling constraints, \scuba
presently has three primary scan modes: PONG (cite Borys et al.),
LISSAJOUS, and DAISY... (more citations? Kovacs et al. has a nice
description... also CMB literature).

@@

Both bolometers share significant long-timescale structure
($\gsim10$\,s) that appears to be related to variations in the fridge
base temperature, although the similarity is clearly greater at
850\,\micron. In this particular case, the total power in the
fluctuations at 450\,\micron, from $-0.04$\,pW to $+0.08$\,pW, are larger
than the $-0.03$\,pW to $+0.05$\,pW fluctuations at 850\,\micron. Such
behaviour might be expected if there is a comparable varying thermal
load from the fridge at each wavelength, but a larger contribution of
atmospheric variations through the 450\,\micron\ bandpass filters. We
also note that there is no obvious strong correlation between the
low-frequency signal structure, at either wavelength, with the
telescope motion.

The low-frequency signal component of the bolometer output is also
highly correlated amongst bolometers in the same subarray. We have
calculated a common-mode signal, $\mathbf{c}(t)$, as the average time
series of all the working bolometers. We then fit the amplitude of
$\mathbf{c}(t)$ at each wavelength to the signals shown in
Fig.~\ref{fig:bolos_mix} and remove it, yielding the grey residual
signals. These residuals are quite flat, although still with
noticeable variations. The white noise is also apparent, and larger at
450\,\micron\ as one would expect from the larger backgrounds compared
to 850\,\micron.

Next, we produce power spectral density (PSD) plots for several
representative bolometers in Fig.~\ref{fig:pspec}. To produce this
figure, we follow the convention that the PSD as a function of
frequency, $\mathbf{P}(f)$, is normalized such that the integral over
frequency gives the same variance as the time-series variance across
the full time series. In other words, given a bolometer signal
$\mathbf{b}(t)$,
%
\begin{equation}
\label{eq:psd}
\langle\mathbf{b}^2(t)\rangle = \int \mathbf{P}(f) df .
\end{equation}
%
However, we only show the PSD up to the Nyquist frequency, so the
conversion from variance measured in the frequency domain to the time
domain requires an additional multiplcation by 2; hence the units are
pW$^2$\,Hz$^{-1}$ rather than pW$^2$\,s.
